---
title: "Allergies"
author: "BABDGroup3"
---

```{r}
library('tm')
library('proxy')
library('ggplot2')
library('wordcloud')
library('textstem')
```

Import data
```{r}
serious <- read.csv(file="2021VAERSDATA+VAX_serious.csv") 
non_serious <- read.csv(file="2021VAERSDATA+VAX_nonserious.csv")
```

***********************
* Serious - ALLERGIES *
***********************

Remove rows with missing ALLERGIES
```{r}
serious <- subset(serious, ALLERGIES=!"")
```

Load SYMPTOM_TEXT as corpus
```{r}
allergy <- VCorpus(VectorSource(serious$ALLERGIES)) 
allergy[[2]]$content
```

----------------------------Text cleaning------------------------------------------------------------------------------

1) Remove punctuation and numbers
```{r}
allergy <- tm_map(allergy,removePunctuation)
allergy <- tm_map(allergy,removeNumbers)
allergy[[2]]$content
```

2) Convert to lowercase
```{r}
allergy <- tm_map(allergy,content_transformer(tolower))
allergy[[2]]$content
```

3) Remove stopwords and strip unnecesary whitespace
```{r}
allergy <- tm_map(allergy, removeWords, stopwords("english"))
allergy <- tm_map(allergy, stripWhitespace)
allergy[[2]]$content
```

Lemmatize using textstem
```{r}
library(textreg)
strings <- convert.tm.to.character(allergy)
```

```{r}
strings_lem <- lemmatize_strings(strings)
```

```{r}
allergy_lem <- VCorpus(VectorSource(strings_lem))
allergy_lem[[2]]$content
```

-------------------------------Document-terms matrix-------------------------------------------------------------------

Create the document-terms matrix
```{r}
dtm <- DocumentTermMatrix(allergy_lem)
dtmmtx <- as.matrix(dtm)
```

Terms and their frequencies
```{r}
freq <- colSums(as.matrix(dtm)) 
wf <- data.frame(word=names(freq),freq=freq)
```

```{r}
dtms <- removeSparseTerms(dtm, 0.99)
freq <- colSums(as.matrix(dtms)) 
wf <- data.frame(word=names(freq),freq=freq)
```

------------------------------Data viz---------------------------------------------------------------------------------

Display the remaining terms according to their frequencies
```{r}
p <- ggplot(subset(wf),aes(word, freq))    
p <- p + geom_bar(stat="identity")   
p <- p + theme(axis.text.x=element_text(angle=90, hjust=1))   
p
```

Display the clouds of words
```{r fig.height=7, fig.width=7}
png('Allergies_Serious.png')
wordcloud(names(freq),freq,scale=c(5,0.1),colors=brewer.pal(6,"Dark2")) 
```

***************************
* Non Serious - ALLERGIES *
***************************

Remove rows with missing ALLERGIES
```{r}
non_serious <- subset(non_serious, ALLERGIES=!"")
```

Load SYMPTOM_TEXT as corpus
```{r}
allergy2 <- VCorpus(VectorSource(non_serious$ALLERGIES)) 
allergy2[[2]]$content
```

----------------------------Text cleaning------------------------------------------------------------------------------

1) Remove punctuation and numbers
```{r}
allergy2 <- tm_map(allergy2,removePunctuation)
allergy2 <- tm_map(allergy2,removeNumbers)
allergy2[[2]]$content
```

2) Convert to lowercase
```{r}
allergy2 <- tm_map(allergy2,content_transformer(tolower))
allergy2[[2]]$content
```

3) Remove stopwords and strip unnecesary whitespace
```{r}
allergy2 <- tm_map(allergy2, removeWords, stopwords("english"))
allergy2 <- tm_map(allergy2, stripWhitespace)
allergy2[[2]]$content
```

Lemmatize using textstem
```{r}
library(textreg)
strings2 <- convert.tm.to.character(allergy2)
```

```{r}
strings_lem2 <- lemmatize_strings(strings2)
```

```{r}
allergy_lem2 <- VCorpus(VectorSource(strings_lem2))
allergy_lem2[[2]]$content
```

-------------------------------Document-terms matrix-------------------------------------------------------------------

Create the document-terms matrix
```{r}
dtm2 <- DocumentTermMatrix(allergy_lem2)
dtmmtx2 <- as.matrix(dtm2)
```

Terms and their frequencies
```{r}
freq2 <- colSums(as.matrix(dtm2)) 
wf2 <- data.frame(word=names(freq2),freq=freq2)
```

```{r}
dtms2 <- removeSparseTerms(dtm2, 0.99)
freq2 <- colSums(as.matrix(dtms2)) 
wf2 <- data.frame(word=names(freq2),freq=freq2)
```

------------------------------Data viz---------------------------------------------------------------------------------

Display the remaining terms according to their frequencies
```{r}
p2 <- ggplot(subset(wf2),aes(word, freq2))    
p2 <- p2 + geom_bar(stat="identity")   
p2 <- p2 + theme(axis.text.x=element_text(angle=90, hjust=1))   
p2
```

Display the clouds of words
```{r fig.height=7, fig.width=7}
png('Allergies_Non_Serious.png')
wordcloud(names(freq2),freq2,scale=c(5,0.1),colors=brewer.pal(6,"Dark2")) 
```































